---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  pdf_document: default
  html_document: default
  word_document: default
---
<center> <h1> Assignment 3 (10%) </h1> </center>
<center>  <h3> Andy Lee </h2> </center>
<center> <h3> DHC 500163559 </h2> </center>
---


#### 1. Import to R the following fiel: http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv (The dataset is related to white Portuguese "Vinho Verde" wine. For more info: https://archive.ics.uci.edu/ml/datasets/Wine+Quality) (3 points)
```{r}
# df <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv", sep = ";")

```

#### 2. Check the datatypes of the attributes. (3 points)
```{r}
# sapply(df, class)
```
  
#### 3. Are there any missing values in the dataset? (4 points)
```{r}
# sum(complete.cases(df)) == nrow(df)
# library(rapportools)
# colSums(!is.empty(df))
# # there appears some missed counts at citric.acid so we further investigate the column.
# df[is.empty(df$citric.acid),]
# Turns out they are just 0s which is normal for citric acid.
# Therefore conclude there are no invalid empty values.
```

#### 4. What is the correlation between the attributes other than Quality? (10 points)
```{r}
# df.noquality <- df[,-12]
# cor(df.noquality)
#plot(df.noquality, pch=20)
```

#### 5. Graph the frequency distribution of wine quality by using Quality. (10 points)
```{r}
# hist(df$quality, main="Wine Quality Distribution", xlab = "Wine Quality")
```

#### 6.	Reduce the levels of rating for quality to three levels as high, medium and low. Assign the levels of 3 and 4 to level 0; 5 and 6 to level 1; and 7,8 and 9 to level 2. (10 points)
```{r}
# df.slicing <- cut(df$quality, breaks=c(2, 4, 6, 9))
# levels(df.slicing) <- c(0:2)
# df$qualitylevels <- df.slicing
```

#### 7. Normalize the data set by using the following function: (12 points)
```{r}
# normalize <- function(x){
#   return ((x - min(x)) / (max(x) - min(x)))
# }
```

```{r}
# df.normalized <- data.frame(lapply(df[1:11], normalize))
# colnames(df.normalized) <- paste(colnames(df.normalized), "norm", sep = ".")
# df.combined <- cbind(df,df.normalized)
```

#### 8. Divide the dataset to training and test sets. (12 points)
```{r}
# set.seed(888)
# train_index <- sample(1:nrow(df.combined), 0.8*nrow(df.combined))
# df.train <- df.combined[train_index,]
# df.test <- df.combined[-train_index,]
```

#### 9. Use the KNN algorithm to predict the quality of wine using its attributes. (12 points)
```{r}
# library(class)
# library(gmodels)
# df.knn <- knn(train = df.train, test = df.test, cl=df.train$qualitylevels, k=3)
```

#### 10. Display the confusion matrix to evaluate the model performance. (12 points)
```{r}
#install.packages('caret', dependencies = TRUE)
# library(caret)
# cm <- confusionMatrix(df.knn, df.test$qualitylevels)
# cm
```

#### 11. Evaluate the model performance by computing Accuracy, Sensitivity and Specificity. (12 points)
```{r}
# writeLines(paste0("Accuracy: ", cm$overall['Accuracy']))
# writeLines("\nSensitivity: TP / (TP + FN)")
# writeLines(paste0("class ", c(0:2),": ", cm$byClass[1:3]))
# writeLines("\nSpecificity: TN / (TN + FP)")
# writeLines(paste0("class ", c(0:2),": ", cm$byClass[4:6]))

```


This is the end of Assignment 3

Ceni Babaoglu, PhD